---
title: "Parenting behaviors"
author: "Emily"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

# Data preprocessing

Preliminaries.

```{r echo=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE)
suppressPackageStartupMessages(c("dplyr","langcog","tidyr","ggplot2","lme4"))
library(psych)
library(langcog)
library(tidyverse)
library(ggthemes)
library(rjson)
library(stringr)
library(forcats)
library(tibble)
library(lme4)
library(lmerTest)

select <- dplyr::select # masked by MASS

theme_set(theme_few())
```

Read in participant data.

```{r}
files <- dir("../production-results/parenting_behaviors_e2/")
attitudes <- data.frame()
behaviors <- data.frame()
subinfo <- data.frame()

for (f in files) {
  jf <- paste("../production-results/parenting_behaviors_e2/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  
  #paq answers
  attitudes_id <- data.frame(sid = as.factor(jd$WorkerId), 
               sent = jd$answers$data$sentence[1:24],
               rating = as.numeric(jd$answers$data$rating[1:24]))
                   
  attitudes <- bind_rows(attitudes, attitudes_id)
  
  #behaviors
  behaviors_id <- data.frame(sid = as.factor(jd$WorkerId),
                sent = jd$answers$data$sentence[25:36],
                rating = as.numeric(jd$answers$data$rating[25:36]))
                          
  behaviors <- bind_rows(behaviors, behaviors_id)
  
  #demographics
  subinfo_id <- data.frame(sid = as.factor(jd$WorkerId),
                   behave_age = jd$answers$data$behaveAge,    
                   children = jd$answers$data$children,
                   youngest = jd$answers$data$childAgeYoung,
                   oldest = jd$answers$data$childAgeOld,
                   ses = jd$answers$data$ladder,
                   gender = jd$answers$data$gender,
                   age = jd$answers$data$age,
                   education = jd$answers$data$education,
                   ethnicity = jd$answers$data$ethnicity,
                   race = as.character(jd$answers$data$race[1]),
                   comments = jd$answers$data$comments)
  subinfo <- bind_rows(subinfo, subinfo_id)
}
```

Read in trial info and questionnaire labels.

```{r}
labels <- read.csv("sent_forms.csv")
labels$sent <- as.character(labels$sent)
behave <- read.csv("behaviors_e2.csv")
```

Clean up labels.

```{r}
attitudes$sent <- str_replace_all(as.character(attitudes$sent), "[â‘”“’']", "")
behaviors$sent <- str_replace_all(as.character(behaviors$sent), "[â‘”“’']", "")
```

Make data frames. 

```{r}
dq <- attitudes %>%
  left_join(labels)%>%
  mutate(category_paq = category)%>%
  select(-category)

db <- behaviors%>%
  left_join(behave)%>%
  left_join(subinfo)%>%
  filter(behave_age != "older")%>%
  mutate(category_bev = category)%>%
  select(-category)

#remove items that parents marked as "my child is too young"
db$rating[db$rating == 0] <- NA

#rescore reverse coded items
dq$rating <- as.numeric(dq$rating)
dq$rating[dq$reverse_code == 1] <- 6 - dq$rating[dq$reverse_code == 1]
```

Get means by category.

```{r}
atts <- dq %>%
  group_by(sid, category_paq) %>% 
  summarise(rating_paq = mean(rating))

bevs <- db %>%
  group_by(sid, category_bev) %>% 
  summarise(rating_bev = mean(rating))

all <- atts %>%
  left_join(bevs)%>%
  left_join(subinfo)%>%
  filter(!children == "0")

all$category_paq <- str_replace(all$category_paq, "AA", "AA_paq")
all$category_paq <- str_replace(all$category_paq, "EL", "EL_paq")
all$category_paq <- str_replace(all$category_paq, "RR", "RR_paq")

all$category_bev <- str_replace(all$category_bev, "AA", "AA_behave")
all$category_bev <- str_replace(all$category_bev, "EL", "EL_behave")
all$category_bev <- str_replace(all$category_bev, "RR", "RR_behave")
```

#Demographics

```{r}
subinfo$education <- factor(subinfo$education, 
                            levels = c("highSchool","someCollege","4year","someGrad","Grad"))

qplot(ses, data=subinfo)
qplot(children, data=subinfo)
qplot(gender, data=subinfo)
qplot(education, data=subinfo)
qplot(age, data=subinfo)
qplot(race, data=subinfo)
qplot(ethnicity, data=subinfo)
```

# PAQ

Get mean ratings for sentences.

```{r}
ms <- dq %>%
  group_by(category_paq, short_sent, reverse_code) %>%
  multi_boot_standard(col = "rating", na.rm = TRUE) %>%
  arrange(category_paq, desc(mean)) 

ms$short_sent_ord <- factor(ms$short_sent, 
                             levels = ms$short_sent)


#for comparison with CDM data
beh <- ms
save(beh, file = "beh.RData")
```

Plot responses to individual questionnaire items.

```{r}
qplot(short_sent_ord, mean, col = category_paq,
      ymin = ci_lower, ymax = ci_upper, pch = factor(reverse_code),
      geom = "pointrange",
      data = ms) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) +
  xlab("") + 
  ylab("Mean Rating") + 
  ylim(c(0,6)) + 
  scale_colour_solarized()
```

Plot mean subscale scores.

```{r}
atts_m <- dq %>%
  group_by(category_paq) %>%
  multi_boot_standard(col = "rating", na.rm = TRUE) %>%
  arrange(category_paq, desc(mean)) 

ggplot(atts_m, aes(x = category_paq, y = mean)) + 
  geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9))+ 
  ylim(c(0,6)) 
```

# Behaviors

Get mean ratings for behaviors.

```{r}
ms_b <- db %>%
  group_by(category_bev, short_sent) %>%
  multi_boot_standard(col = "rating", na.rm = TRUE) %>%
  arrange(category_bev, desc(mean)) 

ms_b$short_sent_ord <- factor(ms_b$short_sent, 
                             levels = ms_b$short_sent)
```

Recode variables for plotting.
```{r}
bev <- behaviors %>%
  left_join(behave)%>%
  filter(rating != 0)%>%
  group_by(category)

bev$rating[bev$rating=="1"] <- "Never"
bev$rating[bev$rating=="2"] <- "Almost never"
bev$rating[bev$rating=="3"] <- "Occasionally"
bev$rating[bev$rating=="4"] <- "Once or twice per week"
bev$rating[bev$rating=="5"] <- "Most days"
bev$rating[bev$rating=="6"] <- "Multiple times ever day"

bev$rating <- factor(bev$rating, levels = c("My child is too young", "Never", "Almost never","Occasionally","Once or twice per week","Most days","Multiple times ever day"))

bev$short_sent <- factor(bev$short_sent, levels = c("read","practice numbers and letters","make observations", "educational programming","talk about feelings","spend time cuddling","sleep in the same bed","hug and kiss","talk sternly","give time out or punishments","talk about setting limits","help with chores"))

subinfo$behave_age[subinfo$behave_age == "0-6"] <- "0-6 months"
subinfo$behave_age[subinfo$behave_age == "7-12"] <- "7-12 months"

subinfo$behave_age <- factor(subinfo$behave_age, levels = c("0-6 months", "7-12 months", "1-1.5","1.5-2","2-2.5","2.5-3","3-3.5","3.5-4", "4-4.5","4.5-5", "older"))
```

Plot mean frequency of individual behavior items. This treats frequency as a continuous variable which it really isn't- is there a better way or does this seem okay?

Parents were asked to answer questions about their youngest child. How old is their youngest child?
```{r}
qplot(behave_age, data=subinfo)
```

How frequent were individual behaviors?
```{r}
qplot(short_sent_ord, mean, col = category_bev,
      ymin = ci_lower, ymax = ci_upper,
      geom = "pointrange",
      data = ms_b) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) +
  xlab("") + 
  ylab("Frequency of Behavior") + 
  ylim(c(0,6)) + 
  scale_colour_solarized()
```

What do the means look like?
```{r}
ggplot(bev, aes(short_sent, fill = rating, shape = category)) + 
  geom_bar(position = "fill")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) +
  xlab("Behavior") + 
  ylab("Proportion responding") 
```


##Reliability

Are behaviors within a category correlated with each other? This is relevant if we choose to use mean scores within these categories for subsequent analyses.

Rules and Respect
```{r}
wide.bev <- db %>% 
  filter(category_bev == "RR") %>%
  select(sid, short_sent, rating) %>% 
  spread(short_sent, rating)
alpha.rr <- as.matrix(select(wide.bev, -sid))
psych::alpha(x = alpha.rr)
```

Affection and Attachment
```{r}
wide.bev <- db %>% 
  filter(category_bev == "AA") %>%
  select(sid, short_sent, rating) %>% 
  spread(short_sent, rating)
alpha.aa <- as.matrix(select(wide.bev, -sid))
psych::alpha(x = alpha.aa)
```

Early Learning
```{r}
wide.bev <- db %>% 
  filter(category_bev == "EL") %>%
  select(sid, short_sent, rating) %>% 
  spread(short_sent, rating)
alpha.el <- as.matrix(select(wide.bev, -sid))
psych::alpha(x = alpha.el)
```

Early Learning and Rules and Respect items hang together pretty well (alphas = .70) but Affection and Attachment items are not as great (alpha = .55). Dropping "sleep in the same bed" would help. Also, "hug and kiss" is close to ceiling with most people reporting "multiple times every day."


##Demographic predictors of behaviors

Education
```{r}
mf <- bevs %>% 
  left_join(subinfo)%>%
  filter(education != "") %>%
  group_by(education, category_bev) %>%
  multi_boot_standard(col = "rating_bev", na.rm = TRUE) 

ggplot(mf, aes(category_bev, mean, fill=education)) +
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) + 
  scale_fill_solarized()
```

Gender
```{r}
mf <- bevs %>% 
  left_join(subinfo)%>%
  filter(gender == "Man" || gender == "Woman") %>%
  group_by(gender, category_bev) %>%
  multi_boot_standard(col = "rating_bev", na.rm = TRUE) 

ggplot(mf, aes(category_bev, mean, fill=gender)) +
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) + 
  scale_fill_solarized()
```
Moms are engaging with their children more overall.

Age of child
```{r}
mf <- bevs %>% 
  left_join(subinfo)%>%
  filter(behave_age != "") %>%
  group_by(behave_age, category_bev) %>%
  multi_boot_standard(col = "rating_bev", na.rm = TRUE) 

ggplot(mf, aes(category_bev, mean, fill=behave_age)) +
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) 
```
More behaviors are possible as kids get older.

Race
```{r}
mf <- bevs %>% 
  left_join(subinfo)%>%
  filter(race != "NULL") %>%
  group_by(race, category_bev) %>%
  multi_boot_standard(col = "rating_bev", na.rm = TRUE) 

ggplot(mf, aes(category_bev, mean, fill=race)) +
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) + 
  scale_fill_solarized()
```

SES
```{r}
mf <- bevs %>% 
  left_join(subinfo)%>%
  filter(ses != "") %>%
  group_by(ses, category_bev) %>%
  multi_boot_standard(col = "rating_bev", na.rm = TRUE) 

ggplot(mf, aes(category_bev, mean, fill=ses)) +
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) 
```

#Attitudes and Behaviors

This might not be the best plot- I'd rather see the individual behaviors, but not sure the best way to present this without overwhelming.

```{r}
ggplot(filter(all, !is.na(category_bev)), aes(x= rating_bev, y = rating_paq)) +
  facet_grid(category_bev ~category_paq) + 
  geom_point(aes(color = category_paq)) +
  xlab("Frequency of behavior") + 
  ylab("PAQ score") + 
  scale_colour_solarized()+
  geom_smooth(method="lm", se=FALSE) 
```

##Stats

Set up dataframes for analyses.
```{r}
d <- db %>%
  left_join(atts)%>%
  left_join(subinfo)%>%
  filter(!is.na(rating))%>%
  filter(children != 0)%>%
  filter(behave_age != "older")%>%
  filter(behave_age != "")%>%
  select(sid, short_sent, rating, category_paq, category_bev, rating_paq, behave_age)%>%
  spread(category_paq, rating_paq)

d$behave_age <- as.character(d$behave_age)

d$behave_age[d$behave_age == "0-6 months"] <- 1
d$behave_age[d$behave_age == "7-12 months"] <- 7
d$behave_age[d$behave_age == "1-1.5"] <- 13
d$behave_age[d$behave_age == "1.5-2"] <- 19 
d$behave_age[d$behave_age == "2-2.5"] <- 25
d$behave_age[d$behave_age == "2.5-3"] <- 31
d$behave_age[d$behave_age == "3-3.5"] <- 37
d$behave_age[d$behave_age == "3.5-4"] <- 43
d$behave_age[d$behave_age == "4-4.5"] <- 49
d$behave_age[d$behave_age == "4.5-5"] <- 55

d$behave_age <- as.numeric(d$behave_age)

d<- d%>%
  mutate(child_age = behave_age)%>%
  select(-behave_age)%>%
  mutate(rating_bin = rating)

d$rating_bin[d$rating %in% 1:4] <- "low"
d$rating_bin[d$rating %in% 5:6] <- "high"

d_aa <- d%>%
  filter(category_bev == "AA")%>%
  filter(!is.na(rating), !is.na(rating_bin))

d_el <- d%>%
  filter(category_bev == "EL")%>%
  filter(!is.na(rating),!is.na(rating_bin))

d_rr <- d%>%
  filter(category_bev == "RR")%>%
  filter(!is.na(rating),!is.na(rating_bin))

d_aa$rating_bin <- as.factor(d_aa$rating_bin)
d_el$rating_bin <- as.factor(d_el$rating_bin)
d_rr$rating_bin <- as.factor(d_rr$rating_bin)
```

Plot

```{r}
d_plot_aa <- d_aa%>%
  gather("subscale", "rating_paq", AA:RR)%>%
  group_by(subscale, rating_bin) %>%
  multi_boot_standard(col = "rating_paq", na.rm = TRUE) 

ggplot(d_plot_aa, aes(subscale, mean, fill=rating_bin)) +
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) 

d_plot_el <- d_el%>%
  gather("subscale", "rating_paq", AA:RR)%>%
  group_by(subscale, rating_bin) %>%
  multi_boot_standard(col = "rating_paq", na.rm = TRUE) 

ggplot(d_plot_el, aes(subscale, mean, fill=rating_bin)) +
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) 

d_plot_rr <- d_rr%>%
  gather("subscale", "rating_paq", AA:RR)%>%
  group_by(subscale, rating_bin) %>%
  multi_boot_standard(col = "rating_paq", na.rm = TRUE) 

ggplot(d_plot_rr, aes(subscale, mean, fill=rating_bin)) +
  geom_bar(stat="identity", position = "dodge") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
             position = position_dodge(width = .9)) 
```

Separate models for each category of behavior. Do PAQ subscale scores predict the frequency of parenting behaviors?

###Affection and Attachment 
```{r}
model <- summary(glmer(rating_bin ~ AA + RR + EL + child_age + 
                (1|sid) + 
                (1|short_sent), #behavior item
              data = d_aa, 
              family = "binomial"))
model
```

###Early Learning 
```{r eval = FALSE}
#for EL behaviors 
model <- summary(glmer(rating_bin ~ AA + RR + EL + child_age + 
                (1|sid) + 
                (1|short_sent), #behavior item
              data = d_el, 
              family = "binomial"))
model
```

###Rules and Respect 
```{r}
model <- summary(glmer(rating_bin ~ AA + RR + EL + child_age + 
                (1|sid) + 
                (1|short_sent), #behavior item
              data = d_rr, 
              family = "binomial"))
model
```

#Conclusions

PAQ scores predict Early Learning and Affection and attachment behaviors in the way we predicted- higher PAQ scores within a subscale predict more frequent parenting behaviors within the same category (i.e., frequency of Affection and Attachment behaviors are predicted by AA PAQ scores but not EL or RR PAQ scores). Additionally, Early Learning and Rules and Respect but not Affection and Attachment behaviors increase with child age.
